Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-10 paper/tv.tex
m

\begin{wrapfigure}{L}{0.39\columnwidth}
  \vspace{-2mm}
\begin{lstlisting}
void addSugar (int n) 
/*require n>=0, _^*
  ensure  t>=n, Sugar#t */
{ if (n == 0) { 
    event ["Sugar"];} 
  else {
    timeout (oneSugar() , 1);
    addSugar (n-1);}} 
\end{lstlisting}
\vspace{-1mm}
\caption{Value-dependent specification.} 
\label{fig:Value_dependent_intro}
\vspace{-1mm}
\end{wrapfigure}

@@

   \code{ \bott
   \  | \ \empt 
    \ | \ \ev
   \  | \ \es_1 {\seq} \es_2
    \ | \ \pi. (\es_1 {\vee} \es_2)
    \ | \ \es_1 {||} \es_2
    \ | \ \pi ?  \es
      %    \  | \ \es_1 \times \es_2
      \ | \ \es \mysharp  t 
   \  | \  \es^\star
}

\\
\m{(Events)} & \ev  &  ::=  & 
\code{
  \anyevent{A}(v, \alpha^*)
    \ | \ \anynotevent{A} 
    \ | \ \_  
}

@@

\begin{figure}[]
  \vspace{-1mm}
  \renewcommand{\arraystretch}{1.2}
\small\begin{align*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%     Disjunction        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \effect_1 \vee \effect_2}  
&\m{iff}\quad & \code{d, s, \varphi \models \effect_1 } \ 
or\ \code{d, s, \varphi \models \effect_2 }  \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Empty          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s,  \varphi \models {\pi} \wedge \empt}  
&\m{iff}\quad & \code{d{=}0}\ and  \ \code{\llbracket {\pi} \rrbracket_s {=}  
{True} } \ and\  \code{ \varphi {=} [] }\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Event          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi} \wedge \ev }  
&\m{iff}\quad & \code{d{=}0}\ and  \ \code{\llbracket {\pi} \rrbracket_s {=}  
{True} }\ and \ \varphi {=} [\ev]  \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%     Concatenate        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi} \wedge (\es_1\cdot \es_2)}  
&\m{iff}\quad & \exists \code{\varphi_1,\varphi_2}. \  
\code{ \varphi_1 {\lappend} \varphi_2 {=} \varphi}\ and\ 
\exists \code{d_1,d_2}. \  
\code{ d_1 {+} d_2 {=} d} \\
&&& \qquad  s.t. \ \code{d_1, s, \varphi_1 {\models} {\pi}  \wedge \es_1 }
\ and \ \code{d_2, s, \varphi_2 {\models} {\pi}  \wedge \es_2 }\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    disjunction         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi} \wedge \pi_1.(\es_1{\vee}\es_2)}  
&\m{iff}\quad & 
\code{\llbracket \pi_1  \rrbracket_{s} {=}  
 {True} },\code{d, s, \varphi \models {\pi}  \wedge \es_1 } 
\ or\ 
\code{\llbracket \pi_1  \rrbracket_{s} {=}  
 {False} },
\code{d, s, \varphi \models {\pi}  \wedge \es_2 } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    parallel         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi} \wedge (\ev_1{\cdot}\es_1)||(\ev_2{\cdot}\es_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models {\pi}  \wedge \ev_1{\cdot} (\es_1||(\ev_2{\cdot}\es_2)) } 
\ or\ \code{d, s, \varphi \models {\pi}  \wedge \ev_2{\cdot}((\ev_1{\cdot}\es_1)||\es_2)  } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    parallel  2        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi} \wedge (\ev\mysharp t_1) || (\epsilon \mysharp  t_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models {(\pi{\wedge}t_1 {\geq}t_2)}, (\ev\mysharp t_1) } 
\ or\ \code{d, s, \varphi \models {(\pi{\wedge}t_1 {<}t_2)}  , (\ev\mysharp t_1){\cdot}(\epsilon \mysharp  (t_2\text{-}t_1))  } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    block         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi} \wedge \pi_1 ? es}  &\m{iff}\quad & 
\code{\llbracket \pi_1  \rrbracket_{s} {=}  
 {True} }, \ \code{d, s, \varphi \models {\pi}  \wedge es}\ or  \ 
  \code{\llbracket \pi_1 \rrbracket_{s} {=}  
{False} }, 
\  \code{d, s, \varphi \models {\pi} \wedge \pi_1 ? es}\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    t times         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi}  \wedge es\mysharp  t}  &\m{iff}\quad &
 \code{\llbracket {\pi}  \wedge t {=} 0 \rrbracket_{s} {=}  
 {True} }, \ \code{d, s, \varphi \models {\pi}  \wedge \empt}\ or  \\
& & &  \code{\llbracket {\pi}\wedge t {>} 0 \rrbracket_{s} {=}  
{True} }, 
\  \exists \code{\es_1,\es_2}. \  \code{\es_1 {\cdot} \es_2 {=} es}\\
&&& \qquad  s.t. \ \code{d, s, \varphi {\models} {(\pi 
{\wedge} t_1{\geq}0{\wedge}t_2{\geq}0 {\wedge} t_1{+} t_2{=}t)}  , 
(\es_1\mysharp t_1) {\cdot} (\es_2\mysharp t_2) } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    star         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi}\wedge es^\star}  &\m{iff}\quad & 
\code{d, s, \varphi \models {\pi}  \wedge \empt}\ or \ 
\code{d, s, \varphi \models {\pi}  \wedge es {\cdot} es^\star}\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    FALSE         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
& \code{d, s, \varphi \models false}  &\m{iff}\quad &
 \code{\llbracket {\pi}  \rrbracket_{s} {=} {False} } \ or \ \code{\varphi {=} \bott}
\end{align*}
  \vspace{-2mm}
\caption{Semantics of \timedEffects.}
\label{fig:Sementic}
\vspace{-2mm}
\end{figure}

>>>>

\begin{wrapfigure}{L}{0.39\columnwidth}
  \vspace{-2mm}
\begin{lstlisting}
void oneSugar() 
/*req: TRUE/\ (_^*).emp
  ens: t>1 /\ emp#t */ 
{ timeout ((), 1); }

void addSugar (int n) 
/*req: n>=0 /\ _^*
  ens: t>=n /\ Sugar#t */
{ if (n == 0) { 
      event ["Sugar"];} 
  else {
      oneSugar();
      addSugar (n-1);}} 
\end{lstlisting}
\vspace{-1mm}
\caption{Value-dependent specification.} 
\label{fig:Value_dependent_intro}
\vspace{-1mm}
\end{wrapfigure}

@@

   \code{ \bott
   \  | \ \empt 
    \ | \ \ev
   \  | \ \es_1 {\seq} \es_2
    \ | \ \es_1 {\vee} \es_2
    \ | \ \es_1 {||} \es_2
    \ | \ \pi ?  \es
      %    \  | \ \es_1 \times \es_2
      \ | \ \es \mysharp  t 
   \  | \  \es^\star
}

\\
\m{(Events)} & \ev  &  ::=  & 
\code{
  \anyevent{A}(v, \alpha^*)
    \ | \ \tau (\pi)
    \ | \ \anynotevent{A} 
    \ | \ \_  
}

@@

\begin{figure}[]
  \vspace{-1mm}
  \renewcommand{\arraystretch}{1.2}
\small\begin{align*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%     Disjunction        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models \effect_1 \vee \effect_2}  
&\m{iff}\quad & \code{d, s, \varphi \models \effect_1 } \ 
or\ \code{d, s, \varphi \models \effect_2 }  \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Empty          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s,  \varphi \models {\pi} \wedge \empt}  
&\m{iff}\quad & \code{d{=}0}\ and  \ \code{\llbracket {\pi} \rrbracket_s {=}  
{True} } \ and\  \code{ \varphi {=} [] }\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%         Event          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi} \wedge \ev }  
&\m{iff}\quad & \code{d{=}0}\ and  \ \code{\llbracket {\pi} \rrbracket_s {=}  
{True} }\ and \ \varphi {=} [\ev]  \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%     Concatenate        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi} \wedge (\es_1\cdot \es_2)}  
&\m{iff}\quad & \exists \code{\varphi_1,\varphi_2}. \  
\code{ \varphi_1 {\lappend} \varphi_2 {=} \varphi}\ and\ 
\exists \code{d_1,d_2}. \  
\code{ d_1 {+} d_2 {=} d} \\
&&& \qquad  s.t. \ \code{d_1, s, \varphi_1 {\models} {\pi}  \wedge \es_1 }
\ and \ \code{d_2, s, \varphi_2 {\models} {\pi}  \wedge \es_2 }\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    disjunction         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi} \wedge \es_1{\vee}\es_2}  
&\m{iff}\quad & 
\code{d, s, \varphi \models {\pi}  \wedge \es_1 } 
\ or\  
\code{d, s, \varphi \models {\pi}  \wedge \es_2 } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    parallel         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi} \wedge (\ev_1{\cdot}\es_1)||(\ev_2{\cdot}\es_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models {\pi}  \wedge \ev_1{\cdot} (\es_1||(\ev_2{\cdot}\es_2)) } 
\ or\ \code{d, s, \varphi \models {\pi}  \wedge \ev_2{\cdot}((\ev_1{\cdot}\es_1)||\es_2)  } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    parallel  2        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi} \wedge (\ev\mysharp t_1) || (\epsilon \mysharp  t_2)}  
&\m{iff}\quad & \code{d, s, \varphi \models {(\pi{\wedge}t_1 {\geq}t_2)}, (\ev\mysharp t_1) } 
\ or\ \code{d, s, \varphi \models {(\pi{\wedge}t_1 {<}t_2)}  , (\ev\mysharp t_1){\cdot}(\epsilon \mysharp  (t_2\text{-}t_1))  } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    block         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi} \wedge \pi_1 ? es}  &\m{iff}\quad & 
\code{\llbracket \pi_1  \rrbracket_{s} {=}  
 {True} }, \ \code{d, s, \varphi \models {\pi}  \wedge es}\ or  \ 
  \code{\llbracket \pi_1 \rrbracket_{s} {=}  
{False} }, 
\  \code{d, s, \varphi \models {\pi} \wedge \pi_1 ? es}\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    t times         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi}  \wedge es\mysharp  t}  &\m{iff}\quad &
 \code{\llbracket {\pi}  \wedge t {=} 0 \rrbracket_{s} {=}  
 {True} }, \ \code{d, s, \varphi \models {\pi}  \wedge \empt}\ or  \\
& & &  \code{\llbracket {\pi}\wedge t {>} 0 \rrbracket_{s} {=}  
{True} }, 
\  \exists \code{\es_1,\es_2}. \  \code{\es_1 {\cdot} \es_2 {=} es}\\
&&& \qquad  s.t. \ \code{d, s, \varphi {\models} {(\pi 
{\wedge} t_1{\geq}0{\wedge}t_2{\geq}0 {\wedge} t_1{+} t_2{=}t)}  , 
(\es_1\mysharp t_1) {\cdot} (\es_2\mysharp t_2) } \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    star         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, s, \varphi \models {\pi}\wedge es^\star}  &\m{iff}\quad & 
\code{d, s, \varphi \models {\pi}  \wedge \empt}\ or \ 
\code{d, s, \varphi \models {\pi}  \wedge es {\cdot} es^\star}\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%    FALSE         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
& \code{d, s, \varphi \models false}  &\m{iff}\quad &
 \code{\llbracket {\pi}  \rrbracket_{s} {=} {False} } \ or \ \code{\varphi {=} \bott}
\end{align*}
  \vspace{-2mm}
\caption{Semantics of \timedEffects.}
\label{fig:Sementic}
\vspace{-2mm}
\end{figure}
